```{r, echo=FALSE, include=FALSE}
honey_sf <- read_sf("datasets/chapter_3/honey.shp")
colnames(honey_sf)[3:8] <- c("Number_of_colonies", "Yield_per_colony",
                          "Total_production", "Stocks","Price_per_lb",
                          "Value_of_production"
)
honey2008 <- honey_sf[honey_sf$year == 2008, ]
state_capitals <- fromJSON("https://raw.githubusercontent.com/vega/vega/master/docs/data/us-state-capitals.json")
state_capitals_sf <- st_as_sf(state_capitals, coords = c("lon", "lat"), crs = 4326)
state_capitals_sf <- state_capitals_sf[!state_capitals_sf$state %in% c("Alaska", "Hawaii"), ]
honey2008_multiline <- st_cast(honey2008, "MULTILINESTRING")
```
# ggplot2
`ggplot2` excels at visualizing all kinds of data and is the "go to package" for most applications, so it should come as no surprise that you can also visualize spatial data with it. For this chapter we will use several extensions to `ggplot2' to create our plots.
```{r, eval=FALSE}
# for loading our data
library(jsonlite)
library(rgdal)
library(sf)
# for plotting
library(extrafont)
library(ggplot2)
library(ggspatial)
library(patchwork)
library(scico)
library(vapoRwave)
# for data wrangling
library(dplyr)
```

## Data used
We won't be using a lot of data in this chapter.
```{r, eval=FALSE}
# load honey shapefile
honey_sf <- read_sf("honey.shp")
# get the data for 2008
honey2008 <- honey_sf[honey_sf$year == 2008, ]
# create a MULTILINESTRING object
honey2008_multiline <- st_cast(honey2008, "MULTILINESTRING")
# load state capitals of the US
state_capitals <- fromJSON(
  "https://raw.githubusercontent.com/vega/vega/master/docs/data/us-state-capitals.json"
  )
# turn it into an sf object
state_capitals_sf <- st_as_sf(state_capitals, coords = c("lon", "lat"), crs = 4326)
# remove alaska and hawaii
state_capitals_sf <- state_capitals_sf[
  !state_capitals_sf$state %in% c("Alaska", "Hawaii"),
  ]
```

## Using ggplot2 to create maps
The easiest way to create spatial plots with ggplot is to use the `geom_sf()` function. By default there is no aesthetic mapping, but we can use arguments like `fill` to easily create choropleth maps.

```{r, fig.align="center", fig.cap="Basic use of ggplot2 for spatial data"}
usa_1 <- ggplot(data = honey2008) +
  geom_sf()
usa_2 <- ggplot(data = honey2008) +
  geom_sf(aes(fill = Price_per_lb)) +
  ggtitle(label = "The Honey crisis of 2008", subtitle = "Price per lb")
usa_1 / usa_2
```

Using the `annotation_*()` functions of `ggspatial` we can customize our maps by adding a base map or other elements to our map.

```{r, fig.align="center", fig.cap="Adding a basemap and some elements",warning=FALSE,message=FALSE}
ggplot(data = honey2008) +
  annotation_map_tile("stamenwatercolor") +
  geom_sf(aes(fill = Price_per_lb), alpha = 0.8) +
  annotation_scale() +
  scale_fill_viridis_c() +
  ggtitle(label = "The Honey crisis of 2008", subtitle = "Price per lb")
```

We can also use the packages `vapoRwave` and `extrafonts`, do some magic and then create some vibey maps.

```{r, fig.align="center",fig.cap= "A E S T H E T I C"}
ggplot(data = honey2008_multiline) +
  geom_sf(color = "#FF6B58", alpha = 0.1,  size = 4) +
  geom_sf(color = "#FF6B58", alpha = 0.1,  size = 3) +
  geom_sf(color = "#FF6B58", alpha = 0.2,  size = 2) +
  geom_sf(color = "#FF6B58", alpha = 0.2,  size = 1) +
  geom_sf(color = "#FF6B58", alpha = 1,  size = 0.5) +
  geom_sf(color = "#F8B660", alpha = 0.1, size = 6, data = state_capitals_sf) +
  geom_sf(color = "#F8B660", alpha = 0.1, size = 5, data = state_capitals_sf) +
  geom_sf(color = "#F8B660", alpha = 0.2, size = 4, data = state_capitals_sf) +
  geom_sf(color = "#F8B660", alpha = 0.2, size = 3, data = state_capitals_sf) +
  geom_sf(color = "#F8B660", alpha = 0.4, size = 2, data = state_capitals_sf) +
  geom_sf(color = "#F8B660", alpha = 1, size = 1, data = state_capitals_sf) +
  labs(subtitle="An aesthetic look",
       title="US States and their Captials", 
       caption = "Excluding Alaska and Hawaii") + 
  new_retro() + 
  scale_colour_newRetro() +
  guides(size = guide_legend(override.aes = list(colour = "#FA5F70FF"))) +
  theme(
    panel.grid.major = element_blank()
  )
```

You can also go completely crazy, as Timo Grossenbacher did in his blog, and create maps like the one below.

```{r, warning=FALSE,message=FALSE,include=FALSE}
theme_map <- function(...) {
  theme_minimal() +
  theme(
    text = element_text(family = "Ubuntu Regular", color = "#22211d"),
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    # panel.grid.minor = element_line(color = "#ebebe5", size = 0.2),
    panel.grid.major = element_line(color = "#ebebe5", size = 0.2),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "#f5f5f2", color = NA), 
    panel.background = element_rect(fill = "#f5f5f2", color = NA), 
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.border = element_blank(),
    ...
  )
}
data <- read.csv("datasets/chapter_2/switzerland/input/avg_age_15.csv", stringsAsFactors = FALSE)
gde_15 <- readOGR("datasets/chapter_2/switzerland/input/geodata/gde-1-1-15.shp", layer = "gde-1-1-15")
# set crs to ch1903/lv03, just to make sure  (EPSG:21781)
crs(gde_15) <- "+proj=somerc +lat_0=46.95240555555556 
+lon_0=7.439583333333333 +k_0=1 +x_0=600000 +y_0=200000 
+ellps=bessel +towgs84=674.374,15.056,405.346,0,0,0,0 +units=m +no_defs"
# fortify, i.e., make ggplot2-compatible
map_data_fortified <- fortify(gde_15, region = "BFS_ID") %>% 
  mutate(id = as.numeric(id))
# now we join the thematic data
map_data <- map_data_fortified %>%
  left_join(data, by = c("id" = "bfs_id"))
# whole municipalities
gde_15_political <- readOGR("datasets/chapter_2/switzerland/input/geodata/g1g15.shp", layer = "g1g15")
crs(gde_15_political) <- "+proj=somerc +lat_0=46.95240555555556 
+lon_0=7.439583333333333 +k_0=1 +x_0=600000 +y_0=200000 
+ellps=bessel +towgs84=674.374,15.056,405.346,0,0,0,0 +units=m +no_defs"
map_data_political_fortified <- fortify(gde_15_political, region = "GMDNR") %>% 
  mutate(id = as.numeric(id))
map_data_political <- map_data_political_fortified %>%
  left_join(data, by = c("id" = "bfs_id"))
map_data_political <- map_data_political[complete.cases(map_data_political),]
# read in background relief
relief <- raster("datasets/chapter_2/switzerland/input/geodata/02-relief-georef-clipped-resampled.tif")
relief_spdf <- as(relief, "SpatialPixelsDataFrame")
# relief is converted to a very simple data frame, 
# just as the fortified municipalities.
# for that we need to convert it to a 
# SpatialPixelsDataFrame first, and then extract its contents 
# using as.data.frame
relief <- as.data.frame(relief_spdf) %>% 
  dplyr::rename(value = `X02.relief.georef.clipped.resampled`)
# remove unnecessary variables
rm(relief_spdf)
rm(gde_15)
rm(map_data_fortified)
rm(map_data_political_fortified)
pretty_breaks <- c(40,42,44,46,48)
# find the extremes
minVal <- min(map_data$avg_age_15, na.rm = TRUE)
maxVal <- max(map_data$avg_age_15, na.rm = TRUE)
# compute labels
labels <- c()
brks <- c(minVal, pretty_breaks, maxVal)
# round the labels (actually, only the extremes)
for(idx in 1:length(brks)){
  labels <- c(labels,round(brks[idx + 1], 2))
}
labels <- labels[1:length(labels)-1]
# define a new variable on the data set just as above
map_data$brks <- cut(
  map_data$avg_age_15, 
  breaks = brks, 
  include.lowest = TRUE,
  labels = labels
  )
brks_scale <- levels(map_data$brks)
labels_scale <- rev(brks_scale)
```

```{r, fig.align="center",fig.cap="Map of Switzerland using ggplot2", warning=FALSE,message=FALSE}
ggplot() +
    # raster comes as the first layer, municipalities on top
    geom_raster(
      data = relief, aes(
        x = x, 
        y = y, 
        alpha = value
        )
      ) +
    # use the "alpha hack"
    scale_alpha(name = "", range = c(0.6, 0), guide = FALSE)  + 
    # municipality polygons
    geom_polygon(
      data = map_data, aes(
        fill = brks, 
        x = long, 
        y = lat, 
        group = group
        )
      ) +
    # municipality outline
    geom_path(
      data = map_data, aes(
        x = long, 
        y = lat, 
        group = group
        ), 
      color = "white",
      size = 0.1
      ) +
    # apart from that, nothing changes
    coord_equal() +
    theme_map() +
    theme(
      legend.position = "bottom",
      plot.title = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5)
      ) +
    labs(
      x = NULL, 
      y = NULL, 
      title = "Switzerland's regional demographics", 
      subtitle = "Average age in Swiss municipalities, 2015", 
      caption = "Geometries: ThemaKart, BFS; Data: BFS, 2016; Relief: swisstopo, 2016"
      ) + 
    scale_fill_manual(
      values = rev(scico(8, palette = "davos")[2:7]),
      breaks = rev(brks_scale),
      name = "Average age",
      drop = FALSE,
      labels = labels_scale,
      guide = guide_legend(
        direction = "horizontal",
        keyheight = unit(2, units = "mm"), 
        keywidth = unit(70/length(labels), units = "mm"),
        title.position = 'top',
        title.hjust = 0.5,
        label.hjust = 1,
        nrow = 1,
        byrow = TRUE,
        reverse = TRUE,
        label.position = "bottom"
      )
    )
```

Overall, `ggplot2` is a great way to easily create geographic maps if you don't want to learn a new plotting package.

[@ggswitz]