# mapdeck
`mapdeck` is an R-package that allows you to create interactive maps using *mapbox GL* and *deck.gl*. Mapbox is a location data platform that can be used to create various applications. Deck.gl is a framework for visualizing large datasets.
To use `mapdeck`, a Mapbox token is required. To obtain a Mapbox token, you must create a free Mapbox account at https://www.mapbox.com. After you have received your token, you can either set it globally with `set_token()` or you can set it manually for each plot with the `token` argument.
```{r, include=FALSE}
library(dotenv)
try(load_dot_env("key.env"), silent = TRUE)
token <- Sys.getenv("MAPBOX_KEY")
```
```{r, eval=FALSE}
# for loading our data
library(sf)
# for datasets
library(nycflights13)
# for plotting
library(mapdeck)
library(RColorBrewer)
# for data wrangling
library(dplyr)
```

## Data used
In this chapter we will use three datasets. The first two are both taken from OpenStreetMap and contain the road network of Hamburg and locations of bakeries in Europe. We will also use a dataset from the package `nycflights13`, which contains all flights from New York airports in 2013.
```{r, eval=FALSE}
roads_hamburg <- read_sf("hamburg.shp")
bakeries <- read_sf("bakeries.kml")
# count the number of flights from each origin to each destination
flights_grouped <- flights %>%
  count(dest, origin)
colnames(airports)[1] <- "dest"
# join the flight data with the coordinates of the destination airports
flights_grouped <- flights_grouped %>%
  right_join(airports[, c(1, 3:4)])
colnames(flights_grouped)[4:5] <- c("lat_dest", "lon_dest")
colnames(airports)[1] <- "origin"
# join the flight data with the coordinates of the origin airports
flights_grouped <- flights_grouped %>%
  right_join(airports[, c(1, 3:4)])
colnames(flights_grouped)[6:7] <- c("lat_origin", "lon_origin")
# remove NAs
flights_grouped <- flights_grouped[!is.na(flights_grouped$dest), ]
```

## Using mapdeck to create maps
If you create a map with mapdeck, you can define a basemap style using `mapdeck_style()`. If you do not define a style, a default mapbox style is used. To create a map, start with `mapdeck()` and then add layers using the pipe operator `%>%`. For instance, we can visualize our flight data by adding an arc layer that represents the flight routes as raised arcs.

```{r,eval=FALSE}
ms <- mapdeck_style("satellite")
mapdeck(data = flights_grouped, style = ms, pitch = 30, token = token) %>%
  add_arc(
    # coordinates of origin airports
    origin = c("lon_origin", "lat_origin"),
    # coordinates of destination airports
    destination = c("lon_dest", "lat_dest"),
    # color our strokes depending on the number of flights
    stroke_from = "n",
    stroke_to = "n",
    palette = "viridis",
    legend = TRUE
    ) %>%
  # set the view
  mapdeck_view(
    location = c(-110, 48),
    # set the zoom level
    zoom = 2,
    # set the pitch angle
    pitch = 45,
  )
```


```{r, fig.align="center", fig.cap="Displaying flights as arcs",echo=FALSE}
htmltools::includeHTML("html_plots/chapter_5/mapdeck_1.html")
```
  

With `add_line()` we can also display our flights as straight lines from origin to destination.

```{r, eval=FALSE}
ms <- mapdeck_style("light")
mapdeck(style = ms, token = token) %>%
  add_line(
    data = flights_grouped,
    origin = c("lon_origin", "lat_origin"),
    destination = c("lon_dest", "lat_dest"),
    stroke_colour = "n",
    stroke_width = 2,
    palette = "plasma",
    legend = TRUE
  ) %>%
  mapdeck_view(
    location = c(-110, 48),
    zoom = 2,
    pitch = 45
  )
```

```{r, fig.align="center", fig.cap="Displaying flights as straight lines", echo=FALSE}
htmltools::includeHTML("html_plots/chapter_5/mapdeck_2.html")
```
  

We can even animate both of these plots with `add_animated_arc()` and `add_animated_line()`.
```{r, eval=FALSE}
ms <- mapdeck_style("outdoors")
mapdeck(style = ms, token = token) %>%
  add_animated_arc(
    data = flights_grouped,
    origin = c("lon_origin", "lat_origin"),
    destination = c("lon_dest", "lat_dest"),
    stroke_from = "n",
    stroke_to = "n",
    stroke_width = 3,
    # set the length of the trails
    trail_length = 10,
    palette = "magenta2green",
    legend = TRUE
  ) %>%
  mapdeck_view(
    location = c(-110, 48),
    zoom = 2,
    pitch = 45
  )
```

```{r, fig.align="center", fig.cap="Displaying flights as animated arcs", echo=FALSE}
htmltools::includeHTML("html_plots/chapter_5/mapdeck_3.html")
```
  

```{r, eval=FALSE}
ms <- mapdeck_style("dark")
mapdeck(style = ms, token = token) %>%
  add_animated_line(
    data = flights_grouped,
    origin = c("lon_origin", "lat_origin"),
    destination = c("lon_dest", "lat_dest"),
    stroke_colour = "n",
    stroke_width = 2,
    # set the speed of the animation
    animation_speed = 10,
    palette = "viridis",
    legend = TRUE
  ) %>%
  mapdeck_view(
    location = c(-110, 48),
    zoom = 2,
    pitch = 45
  )
```
```{r, fig.align="center", fig.cap="Displaying flights as animated lines", echo=FALSE}
htmltools::includeHTML("html_plots/chapter_5/mapdeck_4.html")
```
  

To visualize a `LINESTRING` or `MULTILINESTRING` we can use the `add_path()` function. In the following, the road network of Hamburg is displayed using this function.
```{r, eval=FALSE}
mapdeck(token = token) %>%
  add_path(
    data = roads_hamburg,
    stroke_colour = "#404040"
    ) %>%
  mapdeck_view(
    location = c(10, 53.6),
    zoom = 8
  )
```

```{r, fig.align="center", fig.cap="Displaying the road network of Hamburg", echo=FALSE}
htmltools::includeHTML("html_plots/chapter_5/mapdeck_5.html")
``` 
  

Next, we color our streets depending on the type of street with the argument `stroke_colour`. This time we will also use our custom palette, which we create with `colorRamp()`.

```{r, eval=FALSE}
mapdeck(style = ms, token = token) %>%
  add_path(
    data = roads_hamburg,
    stroke_colour = "type",
    stroke_width = 2,
    legend = TRUE,
    palette = colorRamp(c("#FFB17A", "#B967FF", "#05FFA1"))( (1:256)/256 )
    ) %>%
  mapdeck_view(
    location = c(10, 53.6),
    zoom = 8
  )

```

```{r, fig.align="center", fig.cap="Coloring roads with a custom palette", echo=FALSE}
htmltools::includeHTML("html_plots/chapter_5/mapdeck_6.html")
```
  

Lastly, we will plot spatial `POINT` data, in this case a list of bakeries in Europe. We can plot these as different types of heatmaps using `add_grid()`, `add_hexagon()` or `add_heatmap()`.

```{r, eval=FALSE}
mapdeck(style = ms, token = token) %>%
  add_grid(
    data = bakeries,
    # size of each cell in meters
    cell_size = 1000,
    # multiplier for the elevation of the cells
    elevation_scale = 75,
    colour_range = brewer.pal(6, "YlOrRd")
    ) %>%
  mapdeck_view(
    location = c(10, 48),
    zoom = 3,
    pitch = 45,
  )
```
```{r, fig.align="center", fig.cap="Grid style heatmap of bakeries in Europe", echo=FALSE}
htmltools::includeHTML("html_plots/chapter_5/mapdeck_7.html")
```
  

```{r, eval=FALSE}
mapdeck(style = ms, token = token) %>%
  add_hexagon(
    data = bakeries,
    # radius of each cell in meters
    radius = 1000,
    # multiplier for the elevation of the cells
    elevation_scale = 75,
    colour_range = brewer.pal(6, "YlOrRd")
  ) %>%
  mapdeck_view(
    location = c(10, 48),
    zoom = 3,
    pitch = 45,
  )
```
```{r, fig.align="center", fig.cap="Hexagon style heatmap of bakeries in Europe", echo=FALSE}
htmltools::includeHTML("html_plots/chapter_5/mapdeck_8.html")
```
  

```{r, eval=FALSE}
mapdeck(style = ms, token = token) %>%
  add_heatmap(
    data = bakeries,
    # radius of the circle in pixels
    radius_pixels = 2,
    colour_range = brewer.pal(6, "YlOrRd")
  ) %>%
  mapdeck_view(
    location = c(10, 48),
    zoom = 3,
    pitch = 45,
  )
```
```{r, fig.align="center", fig.cap="Classic heatmap of bakeries in Europe", echo=FALSE}
htmltools::includeHTML("html_plots/chapter_5/mapdeck_9.html")
```
  

The function `add_screengrid()` aggregates points into histogram bins and displays them as a grid. Each time you zoom in or out of the map, the map is re-rendered and the grid is updated.

```{r, eval=FALSE}
mapdeck(style = ms, token = token) %>%
  add_screengrid(
    data = bakeries,
    cell_size = 20,
    opacity = 0.3,
    colour_range = brewer.pal(6, "YlOrRd")
  ) %>%
  mapdeck_view(
    location = c(10, 48),
    zoom = 3,
    pitch = 45,
  )
```
```{r, fig.align="center", fig.cap="Screen grid style heatmap of bakeries in Europe", echo=FALSE}
htmltools::includeHTML("html_plots/chapter_5/mapdeck_10.html")
```
  

## Related links
For more information on the use of `mapdeck` see https://symbolixau.github.io/mapdeck/index.html. [@mapdecklink]