# ggplot2
`ggplot2` excels at visualizing all kinds of data and is the go to package for most applications, so it should come as no surprise, that you can visualize spatial data with `ggplot2` as well. For this chapter we will be using multiple extensions of ggplot2 to create our visualizations.
```{r, eval=FALSE}
library(ggplot2)
library(ggspatial)
library(extrafont)
library(patchwork)
library(vapoRwave)
```

The easiest way to generate spatial plots with ggplot is by using the `geom_spatial()` function. Per default no aesthetic mapping is done, but we can use arguments like `fill` to easily create choropleth maps.

```{r}
usa_1 <- ggplot(data = honey2008) +
  geom_sf()
usa_2 <- ggplot(data = honey2008) +
  geom_sf(aes(fill = Price_per_lb)) +
  ggtitle(label = "The Honey crisis of 2008", subtitle = "Price per lb")
usa_1 / usa_2
```

By using the `annotation_*()` functions from `ggspatial` we can customize our maps by adding a basemap or other elements to our map.

```{r}
ggplot(data = honey2008) +
  annotation_map_tile("stamenwatercolor") +
  geom_sf(aes(fill = Price_per_lb), alpha = 0.8) +
  annotation_scale() +
  scale_fill_viridis_c() +
  ggtitle(label = "The Honey crisis of 2008", subtitle = "Price per lb")
```

```{r, warning=FALSE,message=FALSE,include=FALSE}
theme_map <- function(...) {
  theme_minimal() +
  theme(
    text = element_text(family = "Ubuntu Regular", color = "#22211d"),
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    # panel.grid.minor = element_line(color = "#ebebe5", size = 0.2),
    panel.grid.major = element_line(color = "#ebebe5", size = 0.2),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "#f5f5f2", color = NA), 
    panel.background = element_rect(fill = "#f5f5f2", color = NA), 
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.border = element_blank(),
    ...
  )
}
data <- read.csv("switzerland/input/avg_age_15.csv", stringsAsFactors = FALSE)
gde_15 <- readOGR("switzerland/input/geodata/gde-1-1-15.shp", layer = "gde-1-1-15")
# set crs to ch1903/lv03, just to make sure  (EPSG:21781)
crs(gde_15) <- "+proj=somerc +lat_0=46.95240555555556 
+lon_0=7.439583333333333 +k_0=1 +x_0=600000 +y_0=200000 
+ellps=bessel +towgs84=674.374,15.056,405.346,0,0,0,0 +units=m +no_defs"
# fortify, i.e., make ggplot2-compatible
map_data_fortified <- fortify(gde_15, region = "BFS_ID") %>% 
  mutate(id = as.numeric(id))
# now we join the thematic data
map_data <- map_data_fortified %>% left_join(data, by = c("id" = "bfs_id"))
# whole municipalities
gde_15_political <- readOGR("switzerland/input/geodata/g1g15.shp", layer = "g1g15")
crs(gde_15_political) <- "+proj=somerc +lat_0=46.95240555555556 
+lon_0=7.439583333333333 +k_0=1 +x_0=600000 +y_0=200000 
+ellps=bessel +towgs84=674.374,15.056,405.346,0,0,0,0 +units=m +no_defs"
map_data_political_fortified <- fortify(gde_15_political, region = "GMDNR") %>% 
  mutate(id = as.numeric(id))
map_data_political <- map_data_political_fortified %>% left_join(data, by = c("id" = "bfs_id"))
map_data_political <- map_data_political[complete.cases(map_data_political),]
# read in background relief
relief <- raster("switzerland/input/geodata/02-relief-georef-clipped-resampled.tif")
relief_spdf <- as(relief, "SpatialPixelsDataFrame")
# relief is converted to a very simple data frame, 
# just as the fortified municipalities.
# for that we need to convert it to a 
# SpatialPixelsDataFrame first, and then extract its contents 
# using as.data.frame
relief <- as.data.frame(relief_spdf) %>% 
  dplyr::rename(value = `X02.relief.georef.clipped.resampled`)
# remove unnecessary variables
rm(relief_spdf)
rm(gde_15)
rm(map_data_fortified)
rm(map_data_political_fortified)
pretty_breaks <- c(40,42,44,46,48)
# find the extremes
minVal <- min(map_data$avg_age_15, na.rm = T)
maxVal <- max(map_data$avg_age_15, na.rm = T)
# compute labels
labels <- c()
brks <- c(minVal, pretty_breaks, maxVal)
# round the labels (actually, only the extremes)
for(idx in 1:length(brks)){
  labels <- c(labels,round(brks[idx + 1], 2))
}
labels <- labels[1:length(labels)-1]
# define a new variable on the data set just as above
map_data$brks <- cut(map_data$avg_age_15, 
                     breaks = brks, 
                     include.lowest = TRUE, 
                     labels = labels)
brks_scale <- levels(map_data$brks)
labels_scale <- rev(brks_scale)
```
```{r, fig.align="center",fig.cap="Karte der Schweiz in ggplot2", warning=FALSE,message=FALSE}
ggplot() +
    # raster comes as the first layer, municipalities on top
    geom_raster(data = relief, aes(x = x, 
                                  y = y, 
                                  alpha = value)) +
    # use the "alpha hack"
    scale_alpha(name = "", range = c(0.6, 0), guide = F)  + 
    # municipality polygons
    geom_polygon(data = map_data, aes(fill = brks, 
                                      x = long, 
                                      y = lat, 
                                      group = group)) +
    # municipality outline
    geom_path(data = map_data, aes(x = long, 
                                   y = lat, 
                                   group = group), 
              color = "white", size = 0.1) +
    # apart from that, nothing changes
    coord_equal() +
    theme_map() +
    theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    labs(x = NULL, 
         y = NULL, 
         title = "Switzerland's regional demographics", 
         subtitle = "Average age in Swiss municipalities, 2015", 
         caption = "Geometries: ThemaKart, BFS; Data: BFS, 2016; Relief: swisstopo, 2016") + 
    scale_fill_manual(
      values = rev(scico(8, palette = "davos")[2:7]),
      breaks = rev(brks_scale),
      name = "Average age",
      drop = FALSE,
      labels = labels_scale,
      guide = guide_legend(
        direction = "horizontal",
        keyheight = unit(2, units = "mm"), 
        keywidth = unit(70/length(labels), units = "mm"),
        title.position = 'top',
        title.hjust = 0.5,
        label.hjust = 1,
        nrow = 1,
        byrow = T,
        reverse = T,
        label.position = "bottom"
      )
    )
```

[@ggswitz]