# leaflet
**leaflet** is an open-source JavaScript library that is used to create dynamic online maps. The identically named R package makes it possible to create these kinds of maps in R as well. The syntax is identical to the `mapdeck` syntax. First the function `leaflet()` is called, followed by different layers with `add*()`. Again, the pipe operator `%>%` is used to add layers on top of each other.
```{r, eval=FALSE}
# for loading our data
library(readr)
library(sf)
# for plotting
library(leaflet)
library(leaflet.extras)
# for more stuff
library(dbscan)
library(dplyr)
library(openrouteservice)
library(geosphere)
library(magrittr)
```
## Data used
All data used in this chapter is again taken from OpenStreetMap, and most of the data has been converted to shapefiles.

```{r, eval=FALSE}
pharmacies <- read_sf("pharmacies.shp")
streets <- read_sf("streets.shp")
```
  

## Using leaflet to create maps
In this first example, we will record all pharmacies within a 20-minute travel time window by bicycle from a specific starting point in Munich. First we create our basemap with `leaflet()` and add different provider tiles and a layers control so that users can switch between the different basemaps.
```{r, eval=FALSE}
basemap <- leaflet() %>%
  # add different provider tiles
  addProviderTiles(
    "OpenStreetMap",
    # give the layer a name
    group = "OpenStreetMap"
  ) %>%
  addProviderTiles(
    "Stamen.Toner",
    group = "Stamen.Toner"
  ) %>%
  addProviderTiles(
    "Stamen.Terrain",
    group = "Stamen.Terrain"
  ) %>%
  addProviderTiles(
    "Esri.WorldStreetMap",
    group = "Esri.WorldStreetMap"
  ) %>%
  addProviderTiles(
    "Wikimedia",
    group = "Wikimedia"
  ) %>%
  addProviderTiles(
    "CartoDB.Positron",
    group = "CartoDB.Positron"
  ) %>%
  addProviderTiles(
    "Esri.WorldImagery",
    group = "Esri.WorldImagery"
  ) %>%
# add a layers control
  addLayersControl(
    baseGroups = c(
      "OpenStreetMap", "Stamen.Toner",
      "Stamen.Terrain", "Esri.WorldStreetMap",
      "Wikimedia", "CartoDB.Positron", "Esri.WorldImagery"
    ),
    # position it on the topleft
    position = "topleft"
  )
```
  

Next we add a marker for our starting point. To make it stand out from the markers we will add later, we create a unique looking marker with `makeAwesomeIcon()` and add it to our basemap with `addAwesomeMarkers()`. Notice how we can add layers to our `basemap` object using `%>%`.

```{r, eval=FALSE}
icon.fa <- makeAwesomeIcon(
  icon = "flag", markerColor = "red",
  library = "fa",
  iconColor = "black"
)

map_1 <- basemap %>%
  addAwesomeMarkers(
    lat = 48.1,
    lng = 11.5,
    label = "Starting point",
    icon = icon.fa
  )

map_1
```


```{r, fig.align="center", fig.cap="A simple leaflet map with different layers and one marker", echo=FALSE}
htmltools::tags$iframe(src = "html_plots/chapter_6/leaflet_1.html")
```

Next, we have to calculate the drivetime window from our starting point. For this we will be using the `openrouteservice` package which can be used to calculate drivetime windows from a given point by foot, bike or car for instance. To use the package, you need a token which you can get for free at https://openrouteservice.org/. After calculating the isochrone, we calculate the intersection between the pharmacies in Munich and the drivetime window.

```{r, eval=FALSE}
drivetime <- ors_isochrones(
  # set the starting point
  locations = c(11.5, 48.1),
  # use a cycling profile
  profile = "cycling-regular",
  # 20 minutes drivetime
  range = 1200,
  # return a sf object
  output = "sf",
  # token
  api_key = token
  )

# get the pharmacies within the 20 minutes drivetime
pharmacies_inter <- st_intersection(drivetime, pharmacies)
```
  
We add our drivetime window to the map using `addPolygons()`. Finally, we add our pharmacies using `addMarkers()`. We create custom labels that will show on hover through the `label` argument. To get a grasp of how far away each points is from our starting point, we calculate the distance using `distHaversine()` from the `geosphere` package.
We also add a legend, to show that we have calculated a 20 minute drivetime window.
```{r, eval=FALSE}
pharmacies_inter$distance <- distHaversine(st_coordinates(pharmacies_inter), c(11.5, 48.1))
map_2 <- map_1 %>%
  addPolygons(
    data = drivetime,
    # set the color of the polygon
    color = "#E84A5F",
    # set the opacity of the outline
    opacity = 1,
    # set the stroke width in pixels
    weight = 1,
    # set the fill opacity
    fillOpacity = 0.6
  )

map_3 <- map_2 %>%
  # add pharmacies
  addMarkers(
    data = pharmacies_inter,
    # create custom labels
    label = paste(
      "Name: ", pharmacies_inter$name, "<br>",
      "Distance from location: ",
      round(pharmacies_inter$distance, 1), " meters", "<br>",
      "Street: ", pharmacies_inter$addr_st
    ) %>%
      lapply(htmltools::HTML)
  ) %>%
  # add a legend
  addLegend(
    colors = "#E84A5F",
    labels = "0 - 20 minutes",
    title = "Drivetime",
    opacity = 1, 
    position = "bottomleft"
  )

map_3
```

```{r, fig.align="center", fig.cap="Pharmacies within a 20 minutes drivetime window by bike", echo=FALSE}
htmltools::tags$iframe(src = "html_plots/chapter_6/leaflet_2.html", width = "100%", height = "700vh")
```

For our next example, we will cluster all the pharmacies in Munich using density based clustering and then display them nicely with a `leaflet` map. First, we cluster our points with `dbscan()` from the `dbscan` package. Next, we draw a polygon around the outermost points for each cluster.

```{r, eval=FALSE}
# apply dbscan to the coordinates of the pharmacies
pharmacies_db <- dbscan(st_coordinates(pharmacies), eps = 0.008, minPts = 10)
# add a cluster variable
pharmacies$cluster <- pharmacies_db$cluster
# count the size of each cluster
pharmacies_count <- pharmacies %>%
  dplyr::group_by(cluster) %>%
  dplyr::mutate(count = n())
# split the pharmacies by cluster
pharmacies_split <- split(pharmacies_count, pharmacies_count$cluster)
# get the coordinates for all of the points that are in a cluster
pharmacies_split_coord <- lapply(pharmacies_split[2:length(pharmacies_split)], st_coordinates)
# compute which points lie on the convex hull of each cluster
pharmacies_split_chull <- lapply(pharmacies_split_coord, chull)
# keep only those points
pharmacies_outer <- lapply(seq_len(length(pharmacies_split_chull)), function(x, ...) {
  pharmacies_split_coord[[x]][pharmacies_split_chull[[x]], ]
})
# turn these points into a slightly buffered polygon
pharmacies_outer_sf <- lapply(pharmacies_outer, function(x) {
  # append the last point so that a polygon can be drawn
  x <- rbind(x, x[1, ])
  # turn the points into a polygon
  poly <- st_sfc(st_polygon(list(x))) %>%
    as.data.frame() %>%
    # set the crs system of the points
    st_as_sf(crs = 4326) %>%
    # transform the polygons
    st_transform(3035) %>%
    # buffer the polygons by 200 meters
    st_buffer(200) %>%
    # re-transorm the polygons
    st_transform(4326)
})

# bind the polygons together
clusters <- Reduce(rbind, pharmacies_outer_sf)
# set the count of points in no cluster to NA
pharmacies_count$count[pharmacies_count$cluster == 0] <- NA
```
  

To create our map, we will first define a custom palette using the `colorNumeric()` function, since we will color our points depending on the size of the cluster they are in. Depending on the values that can be mapped to a palette, we can also use the functions `colorBin()`, as well as `colorQuantile()` for numeric variables or `colorFactor()` for factor/character variables.

```{r, eval=FALSE}
# define a custom palette
pal <- colorNumeric(
  c("#E1F5C4", "#EDE574", "#F9D423", "#FC913A", "#FF4E50"),
  # colors depend on the count variable
  domain = pharmacies_count$count,
  )

# corrected html and css to show NA nicely in the legend 
# CSS to correct spacing
css_fix <- "div.info.legend.leaflet-control br {clear: both;}"
# Convert CSS to HTML
html_fix <- htmltools::tags$style(type = "text/css", css_fix)

leaflet() %>%
  # add a dark basemap
  addProviderTiles("CartoDB.DarkMatter") %>%
  # add the polygons of the clusters
  addPolygons(
    data = clusters,
    color = "#E2E2E2",
    # set the opacity of the outline
    opacity = 1,
    # set the stroke width in pixels
    weight = 1,
    # set the fill opacity
    fillOpacity = 0.2
  ) %>%
  # add the pharmacies
  addCircleMarkers(
    data = pharmacies_count,
    # color tthe circles depending on the count
    color = ~pal(count),
    # set the opacity of the circles
    opacity = 0.65,
    # set the radius of the circles
    radius = 4,
    # create custom labels
    label = paste(
      "Name:",
      pharmacies_count$name, "<br>",
      "Cluster:",
      pharmacies_count$cluster, "<br>",
      "Cluster size:",
      pharmacies_count$count, "<br>",
      "Street: ",
      pharmacies_count$addr_st
    ) %>%
      lapply(htmltools::HTML),
  ) %>%
  # add a legend
  addLegend(
    data = pharmacies_count,
    pal = pal,
    values = ~count,
    position = "bottomleft",
    title = "Cluster size:",
    opacity = 0.9
  ) %>%
  # add a minimap
  addMiniMap(tiles = "CartoDB.DarkMatter") %<>%
  # apply the fix
  htmlwidgets::prependContent(html_fix)
```


```{r, fig.align="center", fig.cap="Clustered pharmacies in Munich", echo=FALSE}
htmltools::tags$iframe(src = "html_plots/chapter_6/leaflet_3.html", width = "100%", height = "700vh")
```

An extension of the `leaflet` package is `leaflet.extras`. This package can be used to display heatmaps, for instance. With `addMeasure()` a tape measure is added here to calculate the distance between two points or the area between several points.

```{r, eval=FALSE}
leaflet() %>%
  # add a dark basemap
  addProviderTiles("CartoDB.DarkMatter", group = "CartoDB") %>%
  # add the munich road network
  addPolylines(
    data = streets,
    opacity = 0.5,
    weight = 1,
    color = "white"
  ) %>%
  # add a heatmap
  addWebGLHeatmap(
    data = pharmacies,
    size = 2000,
    units = "m",
    intensity = 0.1,
    gradientTexture = "skyline",
    alphaRange = 1,
    opacity = 0.8
    ) %>%
  # add a measure control to the bottom left
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#0bd3d3",
    completedColor = "#f890e7"
  ) %>%
  addMiniMap("CartoDB.DarkMatter")
```

```{r, fig.align="center", fig.cap="Heatmap of pharmacies in Munich", echo=FALSE}
htmltools::tags$iframe(src = "html_plots/leaflet_4.html", width = "100%", height = "700vh")
```

Since 'leaflet' is a JavaScript library, JavaScript code can also be included. For example, if several background maps and a minimap are added, the first map is always displayed as the minimap even if a different background map is selected. To change this behavior, JavaScript can be included using the function `htmltools::onRender()`.

```{r, eval=FALSE}
basemap %>%
  # add a minimap to our basemap
  addMiniMap(
    # all the tiles in our basemap, display the first one
    tiles = c(
      "OpenStreetMap", "Stamen.Toner", "Stamen.Terrain",
      "Esri.WorldStreetMap", "Wikimedia", "CartoDB.Positron",
      "Esri.WorldImagery"
      )[1],
    toggleDisplay = TRUE) %>%
  # add the pharmacies
  addMarkers(
    data = pharmacies,
    # create custom labels
    popup = paste(
      "Name: ", pharmacies$name, "<br>",
      "Street: ", pharmacies$addr_st
    ) %>%
      lapply(htmltools::HTML)
  ) %>%
  # add javascript code
  htmlwidgets::onRender("
    function(el, x) {
      var myMap = this;
      myMap.on('baselayerchange',
        function (e) {
          myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
        })
    }")
```
```{r, fig.align="center", fig.cap="Updating minimaps when the basemap is updated", echo=FALSE}
htmltools::tags$iframe(src = "html_plots/leaflet_5.html", width = "100%", height = "700vh")
```

JavaScript can also be used to combine point data into clusters that expand when you click on the respective cluster:

```{r, eval=FALSE}
leaflet() %>% 
  addTiles() %>%
  addMarkers(
    data = pharmacies,
    clusterOptions = markerClusterOptions(),
    clusterId = "pointsCluster"
    ) %>%
  addEasyButton(
    easyButton(
      states = list(
        easyButtonState(
          stateName = "unfrozen-markers",
          icon = "ion-toggle",
          title = "Freeze Clusters",
          onClick = JS("
          function(btn, map) {
            var clusterManager =
              map.layerManager.getLayer('cluster', 'pointsCluster');
            clusterManager.freezeAtZoom();
            btn.state('frozen-markers');
          }")
        ),
        easyButtonState(
          stateName = "frozen-markers",
          icon = "ion-toggle-filled",
          title = "UnFreeze Clusters",
          onClick = JS("
          function(btn, map) {
            var clusterManager =
              map.layerManager.getLayer('cluster', 'pointsCluster');
            clusterManager.unfreeze();
            btn.state('unfrozen-markers');
          }")
        )
      )
    )
  )
```

```{r, fig.align="center", fig.cap="Expanding clusters using JavaScript", echo=FALSE}
htmltools::tags$iframe(src = "html_plots/leaflet_6.html", width = "100%", height = "700vh")
```


## Related links
For more information on the use of `leaflet` see https://rstudio.github.io/leaflet/. [@leafletlink]